---
title: "Class Reference"
author: "Devon Milley"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this my personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## To set a working directory

1. Select the folder you want to set as your "home"
2. Click the "more" button in the files panel of RStudio
3. Click "Set as working directory"

## Installing packages

install.packages("name of package")

## Loading libraries

library(name of library)

## To create a code block in a notebook

1. three backticks (the key to the left of the one) + {r}
2. three backticks

## To run code in a code block

Use the green triangle in the top right corner of a code block

## To read an rds file

variable name <- read_rds("file name")

## Glimpse data function

glimpse(variable name)
gives you a sense of the data, some of the values in each coumn, and the data types of each column

## head function

head(variable name)
shows the first six rows of data


## From Pre lab 1
**Task** Run the following codeblock to group the ppp_maryland_loans dataframe by the county for each loan, and then count the number of loans in that group.

%>% = and then

```{r}
ppp_maryland_loans %>% #start with this dataset
  group_by(project_county_name) %>% #group by this variable
  summarise( # summarise function
    count_loans = n() #n() is counting
  )
```

## From Lab 1
**A7** There is a huge range in the loan amounts, but the mean and median loan amounts are much closer to the minimum amount. This probably means that there were a small number of businesses that received the vast amount of the total loan money. 

```{r}
# Write the code to produce your answer here
wv_summary <- wv_ppp_loans %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  )
```
