---
title: "Class Reference"
author: "Devon Milley"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this my personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## To set a working directory

1. Select the folder you want to set as your "home"
2. Click the "more" button in the files panel of RStudio
3. Click "Set as working directory"

## Installing packages

install.packages("name of package")

## Loading libraries

library(name of library)

## To create a code block in a notebook

1. three backticks (the key to the left of the one) + {r}
2. three backticks

## To run code in a code block

Use the green triangle in the top right corner of a code block

## To read an rds file

variable name <- read_rds("file name")

## To read a csv file

variable name <- read_csv("file name")

this command can read a zipped file.
read_csv, is a tidyverse function; read.csv is an R function.

## Glimpse data function

glimpse(variable name)
gives you a sense of the data, some of the values in each coumn, and the data types of each column

## head function

head(variable name)
shows the first six rows of data


## From Pre lab 1
**Task** Run the following codeblock to group the ppp_maryland_loans dataframe by the county for each loan, and then count the number of loans in that group.

%>% = and then

```{r}
ppp_maryland_loans %>% #start with this dataset
  group_by(project_county_name) %>% #group by this variable
  summarise( # summarise function
    count_loans = n() #n() is counting
  )
```

## From Lab 1
**A7** There is a huge range in the loan amounts, but the mean and median loan amounts are much closer to the minimum amount. This probably means that there were a small number of businesses that received the vast amount of the total loan money. 

```{r}
# Write the code to produce your answer here
wv_summary <- wv_ppp_loans %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  )
```

## Summarise
Summarise allows us to choose what parts of the data we want to look at

## Arrange
Arrange allows us to organize how we want to look at the data

## Mutate
Mutate creates new variables/columns

## Select
Select allows us to choose which columns we want to look at

## if_else
prom prelab 2
```{r}
# if/else example
maryland_with_in_out <- maryland_ppp %>%
  mutate(
    in_out = if_else(
        servicing_lender_state == 'MD', "IN", "OUT"
      )
  )
```
if lender state is maryland, assign value of in; if lender state is not maryland, assign value of out

## case_when
from prelab 2
```{r}
# case_when example
maryland_jobs_categories <- maryland_ppp %>%
mutate(
  jobs_retained_category = case_when(
    jobs_retained < 10 ~ 'under_10',
    jobs_retained >= 10 & jobs_retained < 25 ~ '10_to_25',
    jobs_retained >= 25 & jobs_retained < 50 ~ '25_to_50',
    jobs_retained >= 50 & jobs_retained < 100 ~ '50_to_100',
    jobs_retained >= 100 & jobs_retained < 250 ~ '100_to_250',
    jobs_retained >= 250 & jobs_retained < 500 ~ '250_to_500',
    jobs_retained == 500 ~ '500'
  )
)
```
This code creates a new column called jobs_retained_cateogry. Case_when says that when the number of jobs retained falls into a certin category, it puts in the green code into the new column.

## Filtering
filter(column_name operator criteria)

ex. filter(jobs_retained == 5)
filter(jobs_retained <= 20)
>=
!=

to do multiple filters:
& = and (ampersand)
| = or (vertical line)

## changing cases

data_name %>% mutate(project_city = string_to_upper(project_city))
data_name %>% mutate(project_city = string_to_lower(project_city))

to make it permanent, save to a new variable.

## nrow(data)
gives you the number of rows in your data

## guess_max() 

from prelab 3
the guess_max() function — used as an argument when reading in files — uses the first however many rows to set the data type. It's used to look at datasets that have a loooot of rows of data (like a million).
```{r}
ppp_maryland_loans <- read_csv("ppp_loans_md.csv.zip", guess_max=10)

```

## problems()

The problems function checks for problems that occurred when loading data. The number of problems equals the number of rows in the tibble

### Changing a data type
**Task from prelab 3** Run the following codeblock to change the data type of the "id" field from a number (dbl) to a character, and then glimpse the data, to see the change. Add a description of what this code does to your reference notebook.

```{r}
ppp_maryland_loans <- ppp_maryland_loans %>%
  mutate(id = as.character(id))

glimpse(ppp_maryland_loans)
```

